---
title: "Premier League Blog"
author: "mytinyshinys"
date: "March 7, 2016"
output:
  html_document:
    fig_caption: yes
    theme: journal
    toc: true
    toc_depth: 2
    code_folding: hide
css: customBlog.css
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
```

```{r data loading, echo=FALSE, include= FALSE}
# # libraries required
library(shiny)
library(shinydashboard)
library(httr)
library(rvest)
library(XML)
library(doBy) # uses MASS which has a eselect conflict with dplyr - need for sequences
library(dplyr) # this masks select from MASS, filter from stats and intersect etc from base
library(timelineR) # conflict with ggvis on add_axis so added to ggvis currently
library(ggvis)

library(RSQLite)
library(lubridate)
library(stringr)
library(markdown)
library(tidyr)
library(shinyBS)
library(scales)
library(ggplot2)
library(leaflet)
library(rCharts)
library(shinythemes)
library(DT)
library(readr)
library(ggmap)
library(rgdal)
library(choroplethr)
library(choroplethrMaps)
library(taucharts)
library(daff)
library(plotly)
#library(crosstalk) no longer needed as can now use event_data() in plotly
library(explodingboxplotR)
library(beeswarm) # just for test

#library(addins)



positions <- read_csv("data/positions.csv") ##
playerGame <- readRDS("data/playerGame.rds") # reducing columns would help - though might be offset by calcs needed later
playerClub <- readRDS("data/playerClub.rds")
summary <- readRDS("data/summary.rds")
leaders <- readRDS("data/leaders.rds")
standings <- readRDS("data/standings.rds")
allGoalsPlayer <- readRDS("data/allGoalsPlayer.rds")
goals <- readRDS("data/goals.rds")
Goals_team_for <- readRDS("data/Goals_team_for.rds")
Goals_team_ag <- readRDS("data/Goals_team_ag.rds")
Play <- readRDS("data/Play.rds")
Place <- readRDS("data/Place.rds")
Method<- readRDS("data/Method.rds")
teamGames <- readRDS("data/teamGames.rds")
managers <- readRDS("data/managers.rds")
milestones<- read_csv("data/milestones.csv")
playerGeos <- read_csv("data/playerGeos.csv")
goalSeqs <- readRDS("data/goalSeqs.rds")
goalSeqsClub <- readRDS("data/goalSeqsClub.rds")


teamCodes <- teamGames %>% 
  ungroup() %>% 
  select(TEAMNAME,TEAMID) %>% 
  unique()

## playerid name and lat/lon
# pgMini <- playerGame %>%  ## so wil only show those that have made an appearance - but that is prob ok
#   select(PLAYERID,name,city,COUNTRY) %>% 
#   unique() %>% 
#   left_join(playerGeos,by=c("PLAYERID"="playerID")) %>% 
#   mutate(place=paste0(city," ",COUNTRY))


pgMini <- playerGame %>%  ## so wil only show those that have made an appearance - but that is prob ok
  select(PLAYERID,name,city,COUNTRY) %>% 
  unique() %>% 
  
  left_join(playerGeos,by=c("PLAYERID"="playerID")) %>% 
  filter(PLAYERID!="OWNGOAL") %>% 
  mutate(place=ifelse(is.na(city),COUNTRY,paste0(city," ",COUNTRY)))

    
playerChoice <- pgMini$PLAYERID
names(playerChoice) <- pgMini$name

teamsChoice <- sort(unique(playerGame$TEAMNAME))
teamsChoice_2 <- c("All Teams",teamsChoice)

seasonChoice <- sort(unique(playerGame$season), decreasing = TRUE)
seasonChoice_2 <- c("All Seasons",seasonChoice)


countryChoice <- sort(unique(playerGame$COUNTRY))
countryChoice_2 <- c("All Countries",countryChoice)

currentYear <-max(standings$season)
currentRound <-max(standings[standings$season==currentYear,]$tmYrGameOrder)


yrs <-sort(unique(standings$season), decreasing=T)

currentYear <- yrs[1] #[1] "2014/15"
lastYear <- yrs[2]
currentRound <-max(standings[standings$season==currentYear,]$tmYrGameOrder)
## take account of few games
if (currentRound<6) {
  currentValue <- 6
} else {
  currentValue <- currentRound
}
print("currentValue")
print(currentValue)

tmYrs <-standings %>%
  select(season,team) %>%
  unique(.)

hth<-data.frame(standings %>%
                  select(team,OppTeam:gameDate,venue,points,res))
hth$tmYrGameOrder <- NULL
## set up order for sequences (got to be prob after changing standings added team back in)
hth <-data.frame(hth %>%
                   group_by(team,OppTeam) %>%
                   arrange(gameDate) %>%
                   mutate(gameOrder=row_number()))

trueGames <- playerGame %>% 
  filter((START+subOn)>0) %>% 
  group_by(PLAYERID) %>% 
  mutate(trueGameOrder=row_number())

# used in sp_birthplace
allPlayers <- playerGame %>% 
  select(name,PLAYERID,COUNTRY) %>% 
  unique()

## removed as  not sure needed and took ages to load
# standard map data for world
# mapData <- readOGR(dsn=".",
#                    layer = "ne_50m_admin_0_countries", 
#                    encoding = "UTF-8",verbose=FALSE)


## used in year on year change
yrs <- unique(standings$season)
teams <- unique(standings$team)
allSeasonTeams <- data.frame(season=rep(yrs, length(teams)),team=rep(teams, length(yrs)))

```

This is aseries of short articles usually related to apps I develop for the [mytinyshinys site](https://mytinyshinys.shinyapps.io/dashboard/) relating to football/soccer. 

Commented out code can be shown or hidden by toggling buttons


##  <span class="post">Everton - What's up</span>
*8th March 2016*

Everton again flattered to deceive at the weekend - losing 3-2 at home to West Ham after taking a two goal lead

In spite of having some highly-talented young players they languish in the bottom half of the table in spite a positive goal difference of 12

But how rare is this? With a bit of data munging and the use of the plotly package the data can be visualized. Hover points for more information


```{r GD, echo=TRUE}

# The standings data.frame lists the position of each team after each round of matches
# but this year's data needs to be recalculated as not all teams have played the same
# number of games. The goal difference is then extrapolated to a full season


df <-standings %>%  
  group_by(season,team) %>% 
  filter(tmYrGameOrder==max(tmYrGameOrder)) %>% 
  select(team,season,final_Pos,tmYrGameOrder,cumPts,cumGD,cumGF) %>% 
  mutate(avPts=round(cumPts/tmYrGameOrder,2))

pastYears<- df %>% 
  filter(season!="2015/16")

thisYear <- df %>% 
  filter(season=="2015/16") %>% 
  ungroup() %>% 
  mutate(totGD=round(cumGD*38/tmYrGameOrder),totGF=round(cumGF*38/tmYrGameOrder)) %>% 
  arrange(desc(avPts),desc(totGD),desc(totGF),team) %>% 
  mutate(pos=row_number())

# From viewing the data, Everton's record is a bit of an outlier  but certainly no
# more than Manchester City in 2003/4. Let's use the annotation feature to highlight them

EVE <- list(
  x = 16,
  y = 1.36,
  text = "Everton 2015/16",
  xref = "x",
  yref = "y",
  showarrow = TRUE,
  arrowhead = 4,
  ax = 20,
  ay = 20
)

MNC <- list(
  x = 1,
  y = 1.08,
  text = "Man City 2003/04",
  xref = "x",
  yref = "y",
  showarrow = TRUE,
  arrowhead = 4,
  ax = 20,
  ay = 20
)

# We can now plot the chart which combines all completed seasons with the current year (via add_trace)
# The points are infoactive - with further details relating to team, season etc highlighted when hovered

pastYears %>% 
  plot_ly(x=cumGD,y=avPts,mode="markers", hoverinfo = "text", name="Past Yrs",
text = paste(team,"<br>", season,"<br> Pos:",final_Pos,"<br> PPG:",avPts,"<br> GD:",cumGD)) %>%
  add_trace(data=thisYear,x=totGD,y=avPts,mode="markers", name="2015/16",hoverinfo = "text", 
            text = paste(team,"<br>", season,"<br> Pos:",pos,"<br> PPG:",avPts,"<br> GD:",totGD)) %>% 

  layout(hovermode = "closest",
         xaxis=list(title="Goal Difference. 2015/16 extrapolated to 38 game season"),
         yaxis=list(title="Average points per game "),
         title="Comparison of Average Points per Game and Goal Diff in Premier League",
         annotations = list(EVE,MNC)
         )   %>% 
 
  config(displayModeBar = F,showLink = F)

```


###  <span class="blogs">myTinyShiny blogs</span>


- [English Premier League](https://mytinyshinys.shinyapps.io/premierBlog)
- [Other Sports](https://mytinyshinys.shinyapps.io/sportsBlog)
- [Socio-Economic](https://mytinyshinys.shinyapps.io/socioEconomicsBlog)
- [Analytics](https://mytinyshinys.shinyapps.io/analyticsBlog)


###  <span class="blogs">Associated web sites</span>


- [English Premier League](https://mytinyshinys.shinyapps.io/premierLeague)
- [World Soccer](https://mytinyshinys.shinyapps.io/worldSoccer)
- [Dashboard](https://mytinyshinys.shinyapps.io/dashboard)